\thetitle

### 2020-12-27 paper/paper.tex
todo: algorithm, title

% \usepackage{academicons}
% \usepackage{cite}
\usepackage[hidelinks,
plainpages=false,
pdftitle={Reinforcement Learning Methods for Operations Research Applications -- Reloaded},
pdfauthor={Manuel Schneckenreither and Stefan Haeussler},
pdfsubject={Reinforcement Learning Methods for Operations Research Applications -- Reloaded},
pdfkeywords={production planning, order release, machine learning, reinforcement learning,
  operations research},
]{hyperref}
\usepackage{url}
\pagestyle{plain}

@@

% \newcolumntype{x}[1]{>{\centering\arraybackslash\hspace{0pt}}p{#1}}
\title{
  Reinforcement Learning Methods for Operations Research Applications -- Reloaded
}
\author{Manuel Schneckenreither% \inst{1}
  % \orcidID{0000-0002-4812-4665}
  \and Stefan Haeussler% \inst{2}
  % \orcidID{0000-0003-2589-1367}
}

@@

Thus, the action space is defined as \(F = \times_{s=1}^{\size{\States}} A_{s}\), where \(A_{s}\) is
a finite set of possible actions. A \emph{policy} is a sequence \(\pol = (f_{1},f_{2},\ldots)\) of
elements \(f_{t} \in F\). Using the policy \(\pol\) means that if the system is in state \(s\) at
time \(t\) the action \(f_{t}(s)\), i.e.% that is
the \(s\)-th component of \(f_{t}\), is chosen. A stationary policy \(\pol = (f,f,\ldots)\) does not
depend on time. In the sequel we are concerned with stationary policies only. The goal in
reinforcement learning (RL) is to find the best stationary policy \(\pol\) for a given problem,
where the phrase ``best'' will be elaborated in detail below.

@@

where \(0 \leqslant \gamma < 1\) is the discount factor and
\(R_{t}^{\pol}(s) = \E_{\pol}[ r(s_{t},a_{t}) \mid s_{t} = s, a_{t} = a]\) the reward received at
time \(t\) upon starting in state \(s\) by following policy \(\pol\)
\citep{Mahadevan96_AverageRewardReinforcementLearningFoundationsAlgorithmsAndEmpiricalResults}.
\MS[t]{how are the update rules?}

@@

% The aim in the discounted framework is to find an optimal policy \(\polopt_{\gamma}\), which when
% followed, maximises the state value for all states \(s\) as compared to any other policy
% \(\pol_{\gamma}\): \(V_{\gamma}^{\polopt_{\gamma}} - V_{\gamma}^{\pol_{\gamma}} \geqslant 0\). This
% criteria is usually referred to as \(\gamma\)-optimality as the discount factor \(\gamma\) is fixed.
% However, this also means that the actual value set for \(\gamma\) determines the best achievable
% policy. For instance, as can be seen in Figure~\ref{fig:printer} setting \(\gamma\) value to
% \(<0.8027\) the printer-loop is preferred over the mail-loop by the agent, although the mail-loop
% accumulates more reward over time, i.e.\@ selecting the mail-loop is a sub-optimal choice. Observe
% that the average reward received per step equals \(1\) for the printer-loop and \(2\) for the
% mail-loop. Thus, the (Blackwell-)optimal policy is to choose the mail-loop. However, if
% \(\gamma < 3^{-\frac{1}{5}} \approx 0.8027\) an agent using discounted reinforcement learning
% chooses the printer loop.

% \begin{figure}[t!]
%   \centering
%   \input{figures/printer}
%   \caption{\label{fig:printer} A MDP with a single action choice in
%     state \(1\), i.e.\@ two different deterministic policies
%     % , in which the agent can choose between doing the printer-loop or the mail-loop  . The
%     % Observe that the average reward received per step equals \(1\) for the printer-loop and \(2\)
%     % for the mail-loop. Thus, the (Blackwell-)optimal policy is to choose the mail-loop. However,
%     % if \(\gamma < 3^{-\frac{1}{5}} \approx 0.8027\) an agent using discounted reinforcement
%     % learning chooses the printer loop.
%     % %
%     \citep[][adapted]{Mahadevan96_OptimalityCriteriaInReinforcementLearning}.
%   }
% \end{figure}

@@

In the common case of unichain MDPs, in which only a single set of recurrent states exists, the
average reward \(\avgrew^{\pol}(s)\) is equal for all states \(s\)
\citep{Mahadevan96_AverageRewardReinforcementLearningFoundationsAlgorithmsAndEmpiricalResults,Puterman94}.
In the sequel we focus on unichain MDPs in this work and thus may simply refer to it as
\(\avgrew^{\pol}\). A policy that maximises the average reward in every state is called
gain-optimal. % Gain-optimality is the least selective criteria an average reward reinforcement
% learning aims for.

@@

where again \(R_{t}^{\pol}(s)\) is the reward received at time \(t\), starting in state \(s\) and
following policy \(\pol\). Note that the bias values are bounded due to the subtraction of the
average reward. Thus the bias value can be seen as the rewards that additionally sum up in case the
process starts in state \(s\). A policy that is gain-optimal is also bias-optimal, if it maximises
the bias values in every state and compared to every other policy.

@@

\subsection{Optimality Criteria}
\label{subsec:More_Refined_Optimality_Criteria}

The aforementioned and informally introduced notions of gain- and bias-optimality are formalised and
generalised to \(n\)-discount-optimality.

\begin{definition}
  Due to \citet{Veinott69} for MDPs a policy \(\polopt\) is \emph{\(n\)-discount-optimal} for
  \(n=-1,0,1,\ldots\) for all states \(s \in \States\) with discount factor
  \(\gamma\) % and \(V_{\gamma}^{\pol}(s)\) being the value function as defined above
  %
  if and only if
  \begin{align*}
    \lim_{\gamma \to 1}(1-\gamma)^{-n}\ (V_{\gamma}^{\polopt}(s) - V_{\gamma}^{\pol}(s)) \geqslant 0 \tpkt
  \end{align*}
\end{definition}

Only if a policy is \(n\)-discount-optimal for all \(n < m\) it can be \(m\)-discount
optimal \citep{Puterman94,Veinott69}.
%
If a policy is \(\infty\)-discount-optimal then it is said to be Blackwell-optimal
\citep{Blackwell62}.
% weakening the argument!!!
That is, for Blackwell-optimal policies \(\polopt\) there exists a discount factor \(\gammaopt < 1\)
such that \(V_{\gamma}^{\polopt}(s) \geqslant V_{\gamma}^{\pol}(s)\) for all
\(\gamma \geqslant \gammaopt\) and under all policies \(\pol\)
\citep{Mahadevan96_SensitiveDiscountOptimalityUnifyingDiscountedAndAverageRewardReinforcementLearning,Blackwell62}.
Informally that means there exists a discount factor \(<1\) which finds Blackwell-optimal policies.
However, (i) in more complex, i.e.\@ real world, MDPs this value can be arbitrary close to \(1\) and
(ii) the difference in state values may be very small, which (iii) due to the need of state value
function approximation likely causes errors when choosing among different actions in the discounted
framework.

For the case of \(n=-1\) this lead to gain-optimality
\(\avgrew^{\polopt}(s) - \avgrew^{\pol}(s) \geqslant 0\) for all policies \(\pol\) and states \(s\)
\citep{Mahadevan96_AnAveragerewardReinforcementLearningAlgorithmForComputingBiasoptimalPolicies}. In
the case of \(n=0\) it describes bias-optimality with \(V^{\polopt}(s) - V^{\pol}(s) \geqslant 0\)
for all policies \(\pol\) and states \(s\)
\citep{Mahadevan96_AnAveragerewardReinforcementLearningAlgorithmForComputingBiasoptimalPolicies}.
%

In the sequel we present the Laurent series expansion which not only links discounted reinforcement
learning with average reward reinforcement learning, but also connects to \(n\)-discount-optimality.

@@

\cite{MillerVeinott1969} established the link between discounted RL state values
\(V_{\gamma}^{\pol}(s)\) and average reward RL values \(\avgrew^{\pol}(s)\) and \(V^{\pol}(s)\)
using the Laurent series expansion as
\begin{align*}
  V_{\gamma}^{\pol}(s) = \frac{\avgrew^{\pol}(s)}{1-\gamma} + V^{\pol}(s) + e_{\gamma}^{\pol}(s) \tcom
\end{align*}

where the error term \(e_{\gamma}^{\pol}(s)\) exists of infinitely many subterms and
\citet{Puterman94} shows that \(\lim_{\gamma \to 1} e_{\gamma}^{\pol}(s) = 0\). Note how the first
term depending on the average reward \(\avgrew^{\pol}(s)\) converges to infinity as \(\gamma\)
increases.

However, an important insight is the connection between \(n\)-discount-optimality and the Laurent
series expansion of \(V_{\gamma}^{\pol}(s)\). Each addend corresponds to one step in
\(n\)-discount-optimality. That is for \((-1)\)-discount-optimality the average reward must be
maximised.
%
Then for \(0\)-discount-optimality the agent has to choose the highest bias values.
%
Finally \(n\)-discount-optimality for \(n \geqslant 1\) requires to maximise the error term
\(e_{\gamma}^{\pol}(s)\) which only exists as \(\gamma\) is strictly less than \(1\). Thus this term
incorporates the number of expected steps and their corresponding reward on the path to the
Blackwell-optimal policy. That is, it optimises the expected reward collected to reach the Blackwell
optimal policy according to the occurrence on the paths, where shorter paths and those which collect
the rewards sooner are preferred.

Note how the Laurent series expansion dissipates these values in single terms and therefore defines
a divide and conquer methodological approach to reinforcement learning. Thus in average reward
reinforcement learning these terms are learned separately, while in the discounted framework the
combined value for a fixed \(\gamma\)-value is learned.

The addends, where \(n=-1,0,\ldots\) denote the coefficients of the Laurent series expansion,
can be reformulated to following constraint problem with expected reward
\(R^{\pol}(s) = \E_{\pol}[ r(s, a) ] = \sum_{a \in \mathcal{A}(s)}p(a | s) r(s, a)\) as shown by
\cite{MillerVeinott1969} and \citet[p.346]{Puterman94}:

\begin{align}
  \label{eq:constr1}\avgrew^{\pol}(s) - \E[\avgrew^{\pol}(s)] & = 0 &  & \text{for } n = -1 \\
  \label{eq:constr2}\avgrew^{\pol}(s) + V^{\pol}(s) - \E[V^{\pol}(s)] & = R^{\pol}(s) && \text{for } n = 0 \\
  \label{eq:constr3}W^{\pol}_{n-1}(s) + W^{\pol}_{n}(s) - \E[W^{\pol}_{n}(s)] & = 0 && \text{for } n \geqslant 1, \text{where } W_{0}^{\pol}(s) = V^{\pol}(s)
\end{align}

\citet[p.343ff]{Puterman94} shows that due to the given degree of freedom if \(n=-1,0,\ldots,M\)
constraints are satisfying the above conditions for all states \(s\), then only
\(\avgrew^{\pol}(s), V^{\pol}(s), W^{\pol}_{1}, \ldots, W^{\pol}_{M-1}\) are unique, whereas
\(W^{\pol}_{M}\) is offset by the vector \(u\) where for the transition probability matrix \(P\) the
vector \(u\) is characterised by \((I-P)u = 0\). Note that \(u\) is determined by the number of
closed irreducible classes of \(P\), that is for ergodic MDPs \(u\) is determined by a single
constant. Average reward learning is based on the above formulation.

@@

\subsection{Algorithm}
\label{subsec:Algorithm}

\begin{algorithm}[t!]
  \begin{algorithmic}[1]
    \State{}Initialise state \(s_{0}\) and network parameters \(\netTarget, \netWorker\) randomly, set
    an exploration rate \(0 \leqslant \plearn \leqslant \pexp \leqslant 1\), exponential smoothing learning rates
    \(0 < \alpha, \gamma < 1\), and discount factors
    \(0 < \gamma_{0} < \gamma_{1} \leqslant 1\).
    \While{the stopping criterion is not fulfilled}
    \State{}\begin{minipage}[t]{0.9\textwidth} With probability \(\pexp\) choose a random action and
      probability \(1-\pexp\) one that fulfills
      \(\max_{a}\lexeq(\X^{\pol}_{\gamma_{1}}(s_{t},a; \netWorker{}),\X^\pol_ {\gamma_{0}}(s_{t},a; \netWorker{}))\). Let
      \(\isRandAct{t}\) indicate if the action was chosen randomly.
    \end{minipage}
    \State{}\begin{minipage}[t]{0.9\textwidth}
    Carry out action \(a_{t}\), observe reward \(r_{t}\) and resulting state \(s_{t+1}\).
    Store the experience \((s_{t}, \isRandAct{t}, a_{t}, r_{t}, s_{t+1})\) in the experience replay
    memory \(M\).
    \end{minipage}
    \If{a non-random action was chosen or \(\pexp > \plearn\)}
    \begin{align*}
      \avgrew^{\pol}  \gets (1- \alpha) \avgrew^{\pol} + \alpha [r_{t} +
      \max_{a}\X^\pol_{\gamma_1}(s_{t+1},a; \netWorker{}) - \X^\pol_{\gamma_1}(s_{t},a_{t}; \netWorker{})]
    \end{align*}
    \EndIf

@@

The model-free average reward adjusted deep RL algorithm is based on the tabular
version of \citet{schneckenreither2020average} and depicted in Algorithm~\ref{alg:near}.
%
%
%
\MS[t]{describe network initialisation method (Uniform, HeAtAl, etc.)}
%
\MS[t]{Extensions: overestimate rho, rho minimum, Multiple Agents (shared state,  shared rho?),
  init-phase (0.5 for expSmthRewRate, always adapt rho, same rho for all agents): exp smooth reward
  as rho, replay memory actions, n-step (describe replay memory), gradient clipping }
\MS[t]{Side info: Replay memory also stores (disallowed) filtered action indices for optimisation purposes and on
episodic tasks needs to store a bool indicating an episode end. Uses Int8 for state features values.  }

>>>>

\newcommand\theauthor{
  Manuel Schneckenreither% \inst{1}
  % \orcidID{0000-0002-4812-4665}
  \and Stefan Haeussler% \inst{2}
  % \orcidID{0000-0003-2589-1367}
}
\newcommand\thetitle{Average Reward Adjusted Deep Reinforcement Learning: Near-Blackwell-Optimal
  Policies for Order Release}

% \usepackage{academicons}
% \usepackage{cite}
\usepackage[hidelinks,
plainpages=false,
pdftitle={\thetitle{}},
pdfauthor={\theauthor{}},
pdfsubject={\thetitle{}},
pdfkeywords={average reward adjusted reinforcement learning, production planning, order release, machine learning,
  operations research},
]{hyperref}
\usepackage{url}
\pagestyle{plain}

@@

% \newcolumntype{x}[1]{>{\centering\arraybackslash\hspace{0pt}}p{#1}}

\author{\theauthor}
\date{\today}
\title{\thetitle}

@@

Thus, the action space is defined as \(F = \times_{s=1}^{\size{\States}} A_{s}\), where \(A_{s}\) is
a finite set of possible actions. A \emph{policy} is a sequence \(\pol = (f_{1},f_{2},\ldots)\) of
elements \(f_{t} \in F\). Using the policy \(\pol\) means that if the system is in state \(s\) at
time \(t\) the action \(f_{t}(s)\), i.e.% that is
the \(s\)-th component of \(f_{t}\), is chosen. A stationary policy \(\pol = (f,f,\ldots)\) does not
depend on time. In the sequel we are concerned with stationary policies only.

@@

where \(0 \leqslant \gamma < 1\) is the discount factor and
\(R_{t}^{\pol}(s) = \E_{\pol}[ r(s_{t},a_{t}) \mid s_{t} = s, a_{t} = a]\) the reward received at
time \(t\) upon starting in state \(s\) by following policy \(\pol\)
\citep{Mahadevan96_AverageRewardReinforcementLearningFoundationsAlgorithmsAndEmpiricalResults}.
%
The aim in the discounted framework is to find an optimal policy \(\polopt_{\gamma}\), which when
followed, maximises the state value for all states \(s\) as compared to any other policy
\(\pol_{\gamma}\): \(V_{\gamma}^{\polopt_{\gamma}} - V_{\gamma}^{\pol_{\gamma}} \geqslant 0\). This
criteria is usually referred to as \(\gamma\)-optimality as the discount factor \(\gamma\) is fixed.
However, this also means that the actual value set for \(\gamma\) determines the best achievable
policy. For instance, as can be seen in Figure~\ref{fig:printer} setting \(\gamma <0.8027\) the
printer-loop is preferred over the mail-loop, although the mail-loop accumulates more
reward over time, i.e.\@ selecting the mail-loop is a sub-optimal choice. Observe that the average
reward received per step equals \(1\) for the printer-loop and \(2\) for the mail-loop. Thus, the
(Blackwell-)optimal policy is to choose the mail-loop. However, if
\(\gamma < 3^{-\frac{1}{5}} \approx 0.8027\) an agent using discounted reinforcement learning
chooses the printer loop as going the printer loop has a higher evaluation.
%
\begin{figure}[t!]
  \centering
  \input{figures/printer}
  \caption{\label{fig:printer} A MDP with a single action choice in
    state \(1\), i.e.\@ two different deterministic policies
    % , in which the agent can choose between doing the printer-loop or the mail-loop  . The
    % Observe that the average reward received per step equals \(1\) for the printer-loop and \(2\)
    % for the mail-loop. Thus, the (Blackwell-)optimal policy is to choose the mail-loop. However,
    % if \(\gamma < 3^{-\frac{1}{5}} \approx 0.8027\) an agent using discounted reinforcement
    % learning chooses the printer loop.
    % %
    \citep[][adapted]{Mahadevan96_OptimalityCriteriaInReinforcementLearning}.
  }
\end{figure}
%

Therefore, in the seek of a more general optimality criteria for determining the best stationary
policy \(\pol\) for a given problem, we introduce \(n\)-discount-optimality.

% \subsection{Optimality Criteria}
% \label{subsec:More_Refined_Optimality_Criteria}

% The aforementioned and informally introduced notions of gain- and bias-optimality are formalised and
% generalised to \(n\)-discount-optimality as follows.

\begin{definition}[\(n\)-Discount-Optimality]
  Due to \citet{Veinott69} for a MDP a policy \(\polopt\) is \emph{\(n\)-discount-optimal} for
  \(n=-1,0,1,\ldots\) for all states \(s \in \States\) with discount factor
  \(\gamma\) % and \(V_{\gamma}^{\pol}(s)\) being the value function as defined above
  %
  if and only if
  \begin{align*}
    \lim_{\gamma \to 1}(1-\gamma)^{-n}\ (V_{\gamma}^{\polopt}(s) - V_{\gamma}^{\pol}(s)) \geqslant 0 \tpkt
  \end{align*}
\end{definition}

Only if a policy is \(n\)-discount-optimal for all \(n < m\) it can be \(m\)-discount
optimal \citep{Puterman94,Veinott69}.
%
If a policy is \(\infty\)-discount-optimal then it is said to be Blackwell-optimal
\citep{Blackwell62}.
% weakening the argument!!!
That is, for Blackwell-optimal policies \(\polopt\) there exists a discount factor \(\gammaopt < 1\)
such that \(V_{\gamma}^{\polopt}(s) \geqslant V_{\gamma}^{\pol}(s)\) for all
\(\gamma \geqslant \gammaopt\) and under all policies \(\pol\)
\citep{Mahadevan96_SensitiveDiscountOptimalityUnifyingDiscountedAndAverageRewardReinforcementLearning,Blackwell62}.
Informally that means there exists a discount factor \(\gamma <1\) which finds Blackwell-optimal
policies. However, (i) in more complex, i.e.\@ real world, MDPs this value can be arbitrary close to
\(1\) and (ii) the difference in state values may be very small, which (iii) due to the need of
state value function approximation likely causes errors when choosing among different actions in the
discounted framework.

% In the sequel we present the Laurent series expansion which not only links discounted reinforcement
% learning with average reward reinforcement learning, but also connects to \(n\)-discount-optimality.

@@

@@

In the common case of unichain MDPs, in which only a single set of recurrent states exists, the
average reward \(\avgrew^{\pol}(s)\) is equal for all states \(s\)
\citep{Mahadevan96_AverageRewardReinforcementLearningFoundationsAlgorithmsAndEmpiricalResults,Puterman94}.
In the sequel we focus on unichain MDPs in this work and thus may simply refer to it as
\(\avgrew^{\pol}\). A policy \(\polopt\) that maximises the average reward
\(\avgrew^{\polopt}(s) - \avgrew^{\pol}(s) \geqslant 0\) in every state \(s\) as compared to any
other policy \(\pol\) is called gain-optimal.
%
% A policy that maximises the average reward in every state is called
% gain-optimal
% \citep{Mahadevan96_AnAveragerewardReinforcementLearningAlgorithmForComputingBiasoptimalPolicies}.
% Gain-optimality is the least selective criteria an average reward reinforcement
% learning aims for.
\citet{Mahadevan96_AnAveragerewardReinforcementLearningAlgorithmForComputingBiasoptimalPolicies}
shows that \(n\)-discount-optimality for \(n=-1\) describes gain-optimality.

@@

where again \(R_{t}^{\pol}(s)\) is the reward received at time \(t\), starting in state \(s\) and
following policy \(\pol\). Note that the bias values are bounded due to the subtraction of the
average reward. Thus the bias value can be seen as the rewards that additionally sum up in case the
process starts in state \(s\). A policy \(\polopt\) that is gain-optimal is also bias-optimal, if it
maximises the bias values \(V^{\polopt}(s) - V^{\pol}(s) \geqslant 0\) in every state \(s\) and
compared to every other policy \(\pol\). Bias-optimality is given by setting \(n=0\) in the
definition of \(n\)-discount-optimality
\citep{Mahadevan96_AnAveragerewardReinforcementLearningAlgorithmForComputingBiasoptimalPolicies}.

@@

@@

\cite{MillerVeinott1969} established the link between discounted RL state values
\(V_{\gamma}^{\pol}(s)\) and average reward RL values \(\avgrew^{\pol}(s)\) and \(V^{\pol}(s)\)
using the Laurent series expansion as
\begin{align*}
  V_{\gamma}^{\pol}(s) = \frac{\avgrew^{\pol}(s)}{1-\gamma} + V^{\pol}(s) + e_{\gamma}^{\pol}(s) \tcom
\end{align*}
%
where the error term \(e_{\gamma}^{\pol}(s)\) exists of infinitely many subterms and
\citet{Puterman94} shows that \(\lim_{\gamma \to 1} e_{\gamma}^{\pol}(s) = 0\). Note how the first
term depending on the average reward \(\avgrew^{\pol}(s)\) converges to infinity as \(\gamma\)
increases.

However, an important insight is the connection between \(n\)-discount-optimality and the Laurent
series expansion of \(V_{\gamma}^{\pol}(s)\). Each addend corresponds to one step in
\(n\)-discount-optimality. That is for \((-1)\)-discount-optimality the addend describes (a scaled
version of) the term to maximise for gain-optimality,
% the average reward must be
% maximised.
%
for \(0\)-discount-optimality the term to maximise for bias-optimality, and
%
finally \(n\)-discount-optimality for \(n \geqslant 1\) requires to maximise the error term
\(e_{\gamma}^{\pol}(s)\). The later only exists as \(\gamma\) is strictly less than \(1\). Thus this
term incorporates the number of expected steps and their corresponding reward on the path to the
Blackwell-optimal policy. Put differently, it optimises the expected reward collected to reach the
Blackwell optimal policy according to the occurrence on the paths, where shorter paths and those
which collect the rewards sooner are preferred.

% Note how the Laurent series expansion dissipates these values in single terms and therefore defines
% a divide and conquer methodological approach to reinforcement learning. Thus in average reward
% reinforcement learning these terms are learned separately, while in the discounted framework a
% single scalar value for each state with a fixed \(\gamma\)-value is estimated.

These addends, where \(n=-1,0,\ldots\) denote the coefficients of the Laurent series expansion and
thus correspond to the \(n\) of the definition of \(n\)-discount-optimality, can be reformulated to
following constraint problem \citep[p.346]{MillerVeinott1969,Puterman94}.
%
\begin{align}
  \label{eq:constr1}\avgrew^{\pol}(s) - \E[\avgrew^{\pol}(s)] & = 0 &  & \text{for } n = -1 \\
  \label{eq:constr2}\avgrew^{\pol}(s) + V^{\pol}(s) - \E[V^{\pol}(s)] & = R^{\pol}(s) && \text{for } n = 0 \\
  \label{eq:constr3}W^{\pol}_{n-1}(s) + W^{\pol}_{n}(s) - \E[W^{\pol}_{n}(s)] & = 0 && \text{for } n \geqslant 1, \text{where } W_{0}^{\pol}(s) = V^{\pol}(s)
\end{align}
%
The error term \(e_{\gamma}^{\pol}(s)\) is given by
\(e_{\gamma}^{\pol}(s) = \sum_{n=1}^{\infty}W^{\pol}_{n}(s)\) and the expected reward
\(R^{\pol}(s) = \E_{\pol}[ r(s, a) ]\). % = \sum_{a \in \mathcal{A}(s)}p(a | s) r(s, a)\)
%
\citet[p.343ff]{Puterman94} shows that due to the given degree of freedom if \(n=-1,0,\ldots,M\)
constraints are satisfying the above conditions for all states \(s\), then only
\(\avgrew^{\pol}(s), V^{\pol}(s), W^{\pol}_{1}, \ldots, W^{\pol}_{M-1}\) are unique, whereas
\(W^{\pol}_{M}\) is offset by the vector \(u\) where for the transition probability matrix \(P\) the
vector \(u\) is characterised by \((I-P)u = 0\). Note that \(u\) is determined by the number of
closed irreducible classes of \(P\), that is for ergodic MDPs \(u\) is determined by a single
constant. Average reward learning is based on the above formulation.

@@

\subsection{Average Reward Adjusted Reinforcement Learning}
\label{subsec:Average_Reward_Adjusted_Reinforcement_Learning}

In the sequel we briefly present the needed formalism of average reward adjusted RL, where for a
more comprehensive version see \citep{schneckenreither2020average}. \MS[t]{todo}

\begin{definition}
  Due to \citet{schneckenreither2020average} the \textit{average reward adjusted (discounted) state
    value} $\X_{\gamma}^{\pol}(s)$ of a state $s$ under policy $\pol$ and with discount factor
  $0 \leqslant \gamma \leqslant 1$ is given as
  \(\X_{\gamma}^{\pol}(s) \defsym V^{\pol}(s) + e_{\gamma}^{\pol}(s)\).
\end{definition}
%
With the Laurent Series expensive this reformulates to
\(\X_{\gamma}^{\pol}(s) = V_{\gamma}^{\pol}(s) - \frac{\avgrew^{\pol}}{1-\gamma}\). Therefore,
\(X_{\gamma}^{\pol}(s)\) describes the discounted state value adjusted by the average reward term.
The Bellman Equation is given by
\(\X_{\gamma}^{\pol}(s) = \E_{\pol}[r_{t} + \gamma \X_{\gamma}^{\pol}(s_{t+1}) - \avgrew^{\pol}
\mid s_{t} = s]\). By turning this formula into an update rule it builds the foundation of our
algorithm.

% \subsection{Algorithm}
% \label{subsec:Algorithm}

\begin{algorithm}[t!]
  \begin{algorithmic}[1]
    \State{}Initialise state \(s_{0}\) and network parameters \(\netTarget, \netWorker\) randomly, set
    an exploration rate \(0 \leqslant \plearn \leqslant \pexp \leqslant 1\), exponential smoothing learning rates
    \(0 < \alpha, \gamma < 1\), and discount factors
    \(0 < \gamma_{0} < \gamma_{1} \leqslant 1\), where \(\gamma_{1} = 1\) is
    usually a good choice.
    \While{the stopping criterion is not fulfilled}
    \State{}\begin{minipage}[t]{0.9\textwidth} With probability \(\pexp\) choose a random action and
      probability \(1-\pexp\) one that fulfills
      \(\max_{a}\lexeq(\X^{\pol}_{\gamma_{1}}(s_{t},a; \netWorker{}),\X^\pol_ {\gamma_{0}}(s_{t},a; \netWorker{}))\). Let
      \(\isRandAct{t}\) indicate if the action was chosen randomly.
    \end{minipage}
    \State{}\begin{minipage}[t]{0.9\textwidth}
    Carry out action \(a_{t}\), observe reward \(r_{t}\) and resulting state \(s_{t+1}\).
    Store the experience \((s_{t}, \isRandAct{t}, a_{t}, r_{t}, s_{t+1})\) in the experience replay
    memory \(M\).
    \end{minipage}
    \If{a non-random action was chosen or \(\pexp > \plearn\)}
    \begin{align*}
      \avgrew^{\pol}  \gets (1- \alpha) \avgrew^{\pol} + \alpha [r_{t} +
      \max_{a}\X^\pol_{\gamma_1}(s_{t+1},a; \netWorker{}) - \X^\pol_{\gamma_1}(s_{t},a_{t}; \netWorker{})]
    \end{align*}
    \EndIf

@@

\paragraph{Average Reward Adjusted Deep RL Algorithm.} %
The model-free average reward adjusted RL algorithm is based on the tabular version of
\citet{schneckenreither2020average} and depicted in Algorithm~\ref{alg:near}.
%
%
%
\MS[t]{describe network initialisation method (Uniform, HeAtAl, etc.)}
%
\MS[t]{Extensions: overestimate rho, rho minimum, Multiple Agents (shared state,  shared rho?),
  init-phase (0.5 for expSmthRewRate, always adapt rho, same rho for all agents): exp smooth reward
  as rho, replay memory actions, n-step (describe replay memory), gradient clipping }
\MS[t]{Side info: Replay memory also stores (disallowed) filtered action indices for optimisation purposes and on
episodic tasks needs to store a bool indicating an episode end. Uses Int8 for state features values.  }
